---
title: "Machine Learning Course Project"
author: "AMR"
date: "December 14, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#loading the packages
library(caret)
library(readr)
library(randomForest)
library(rattle)
library(rpart)
library(rpart.plot)
library(repmis)

```

## Executive summary

The goal of your project is to predict the manner in which the participants did their exercise. This is the "classe" variable in the training set. 
The following variables are used to predict the the manner in which the participants did their exercise:  

This report describe how the prediction model was built, how cross validation was used.

In addition this report describe what needed to be expected out of sample error.

The prediction model will be used to predict 20 different test cases.



## Backgroung

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

## Loading the data

pml_training <- read_csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")
View(pml_training)
str(pml_training)

pml_testing <- read_csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv")
View(pml_testing)



## Cleaning the data


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
